---
title: "Tidy Tuesday 2023 - January - Art History"
output: html_notebook
---

Load relevant packages

```{r}
library(tidyverse)
library(ggplot2)
library(magrittr)

```

Load dataset

```{r}
artists <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-17/artists.csv')

```

Okay lets find out what columns are here
```{r}
str(artists)
```
Okay, we have stuff related to the artist such as; nationality, name, geneder, race, ethnicity, etc; We also have information on which book it came from Janson or Gardner, the amount of space that particular artist was given in the book, the number of exhibitions that a particular artist had in the Moma at time of publication, or the whitney

So what kind of questions can we ask from this data?
First one! How many unique artist are in this dataset?

```{r}

unique_artists <- artists %>%
  select(artist_name, artist_unique_id) %>%
  mutate(artist_id_pair = paste(artist_name, artist_unique_id,sep = "_")) %>%
  summarise(num_of_artists = n_distinct(artist_id_pair))

unique_artists
        
```

Okay by book?
```{r}
unique_artists_by_book <- artists %>%
  select(artist_name, book, artist_unique_id) %>%
  group_by(book) %>%
  mutate(artist_id_pair = paste(artist_name, artist_unique_id,sep = "_")) %>%
  summarise(num_of_artists = n_distinct(artist_id_pair))

unique_artists_by_book
```
WHAT KIND OF BARPLOT CAN WE MAKE?
```{r}
unique_artists_by_year <- artists %>%
  group_by(year, book) %>%
  select(artist_name, artist_unique_id, book, year) %>%
  mutate(artist_id_pair = paste(artist_name, artist_unique_id,sep = "_")) %>%
  arrange(book,year) %>%
  #mutate(year = as.factor(year)) %>%
  summarise(num_of_artists = n_distinct(artist_id_pair))

unique_artists_by_year
```

```{r}
# Gardner - 	#e07261
# Janson - 	#86abc5
#levels(unique_artists_by_year$year) 
unique_artists_by_year %>% ggplot(aes(x = factor(year), y = num_of_artists, fill= factor(book))) +
  theme_bw() +
  theme(panel.grid.major.y = element_blank() ) +
  theme(legend.position = c(0.92, 0.88)) +
  geom_bar(stat='identity', width = 0.7, position = position_dodge(preserve = "single")) +
  scale_fill_manual(drop=FALSE, name = "Textbook",values = c("#e07261", "#86abc5")) +
  scale_x_discrete(limits=rev) +
  coord_flip() +
  labs(y = "Number of Unique Artists", x = "Publication Year", title = "Artist Covered in Major Art History Textbooks From 1926-2020", caption = "based on data from the arthistory data package" ) 
```

```{r}

#levels(unique_artists_by_year$year) 
unique_artists_by_year %>% ggplot(aes(x = year, y = num_of_artists, fill= factor(book))) +
  geom_bar(stat='identity', position = position_dodge(preserve = "single")) +
  scale_fill_discrete(drop=FALSE) +
  coord_flip()
```